---
title: "Data & downloads"
subtitle: "The full dataset, in machine-readable form."
---

## Live Google Sheet

The canonical source of truth is the shared Google Sheet:

> [Open the live Google Sheet →](https://docs.google.com/spreadsheets/d/REPLACE_WITH_YOUR_SHEET_ID/edit?usp=sharing)

This site is rebuilt regularly from this sheet. If you intend to use the data for research, please cite the sheet by its date of access (see [Methods](methods.qmd) for the suggested citation).

## CSV downloads

Snapshots of the two main tabs, fetched at the time this page was last built:

- [`characteristic_data.csv`](data/characteristic_data.csv) — country-level burden, diagnostic, treatment, and implementation indicators.
- [`case_management.csv`](data/case_management.csv) — detailed case management policy data, with sub-rows for primaquine regimen by G6PD status.

```{r}
#| include: false

source("R/data_helpers.R")
char <- load_characteristic_data()
```

The characteristic data file contains **`r ncol(char)`** columns across **`r nrow(char)`** countries.

## Column dictionary

```{r}
#| output: asis

knitr::kable(
  column_dictionary_table(char),
  col.names = c("Column", "Description"),
  format = "pipe"
)
```

## API

There is no separate API; the Google Sheet itself is the API. You can pull either tab as CSV using the standard Sheets endpoint:

```bash
curl "https://docs.google.com/spreadsheets/d/<SHEET_ID>/gviz/tq?tqx=out:csv&sheet=Characteristic+Data" \
  > characteristic_data.csv
```

## Licence

The data is published under the **Creative Commons Attribution 4.0 International (CC-BY-4.0)** licence. You are free to share and adapt the material, including commercially, as long as you give appropriate credit (see citation in [the homepage footer](index.qmd)).
