---
title: "Explore the data"
subtitle: "Filter, search, and download country-level *P. vivax* policy and implementation data."
page-layout: full
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(jsonlite)
  library(crosstalk)
  library(DT)
  library(htmltools)
})

source("R/explore_helpers.R")

char <- load_characteristic_data()
case <- load_case_data(char = char)

char <- char |>
  mutate(
    first_line = collapse_first_line(`Pv 1st line treatment`),
    `2023 Case Numbers` = suppressWarnings(as.numeric(`2023 Case Numbers`))
  )

table_cols <- c(
  "Country", "Region", "WHO Region",
  "Reporting vivax cases (<5 years)",
  "2023 Case Numbers",
  "Proportion of P. vivax cases (2023) (%)",
  "Pv 1st line treatment",
  "Pv 2nd line treatment",
  "Guidelines G6PD testing  (Y/N)",
  "Implementation: G6PD testing  (Y/N)",
  "Type of G6PD testing",
  "Program Phase",
  "Economic status"
)

explore_payload <- explore_chart_payload(char, case)

shared <- SharedData$new(char, key = ~Country, group = "explore")
char_table <- char |> select(any_of(table_cols))
shared_table <- SharedData$new(char_table, key = ~Country, group = "explore")
```

```{r}
#| label: explore-data
#| echo: false
#| results: asis

cat(as.character(htmltools::tags$script(
  id = "explore-data",
  type = "application/json",
  HTML(jsonlite::toJSON(explore_payload, auto_unbox = TRUE, dataframe = "rows", null = "null"))
)))
```

## Filter the data

Use any control below — selections apply to **every chart and the table** on this page.

```{r}
#| label: filters
#| echo: false

tagList(
  tags$div(
    class = "explore-filters",
    bscols(
      filter_select("region", "Region", shared, ~Region, multiple = TRUE),
      filter_select("who_region", "WHO region", shared, ~`WHO Region`, multiple = TRUE),
      filter_select(
        "reporting",
        "Reporting P. vivax (5 yrs)",
        shared,
        ~`Reporting vivax cases (<5 years)`,
        multiple = TRUE
      ),
      filter_select("program_phase", "Programme phase", shared, ~`Program Phase`, multiple = TRUE),
      filter_select(
        "g6pd_guidelines",
        "G6PD testing in guidelines",
        shared,
        ~`Guidelines G6PD testing  (Y/N)`,
        multiple = TRUE
      ),
      filter_select(
        "g6pd_implementation",
        "G6PD testing implemented",
        shared,
        ~`Implementation: G6PD testing  (Y/N)`,
        multiple = TRUE
      ),
      filter_select(
        "first_line",
        "First-line treatment",
        shared,
        ~`Pv 1st line treatment`,
        multiple = TRUE
      ),
      widths = c(4, 4, 4, 4, 4, 4, 4),
      device = "sm"
    )
  )
)
```

## At a glance

::: {.explore-chart}
<div id="chart-cases" class="explore-chart__plot vivax-chart__plot--cases"></div>
<p class="explore-chart__caption">The 10 countries with the highest reported *P. vivax* case burden in the latest available year. This ranking updates with the filters above.</p>
:::

## Repository-wide and regional *P. vivax* case trends

::: {.explore-chart}
<div id="chart-case-trends" class="explore-chart__plot vivax-chart__plot--case-trends"></div>
<p class="explore-chart__caption">Regional and repository-wide totals use a comparable panel of countries with a reported value in every available year. The lower panel shows the average per country using that same panel.</p>
:::

## Regional *P. vivax* cases: latest comparable years

::: {.explore-chart}
<div id="chart-regional-cases" class="explore-chart__plot vivax-chart__plot--regional-cases"></div>
<p class="explore-chart__caption">Uses the two latest annual case columns available and includes only countries reporting values in both years. It will update automatically when newer annual columns are added.</p>
:::

## First-line *P. vivax* treatment policy ranking

::: {.explore-chart}
<div id="chart-treatment-ranking" class="explore-chart__plot vivax-chart__plot--treatment-ranking"></div>
<p class="explore-chart__caption">Number of countries reporting each recommended first-line treatment. This represents national policy recommendations, not medicine consumption or patient-level use.</p>
:::

## Countries with the oldest reported policy updates

::: {.explore-chart}
<div id="chart-oldest-policy" class="explore-chart__plot vivax-chart__plot--oldest-policy"></div>
<p class="explore-chart__caption">The 10 oldest known national policy-update years among the selected countries. Countries without a reported year are excluded rather than treated as having an old policy.</p>
:::

## Full data table

The table below uses the same filters as the charts. To download the complete dataset (all columns), see [Data & downloads](data.qmd).

```{r}
#| label: table

datatable(
  shared_table,
  filter = "none",
  selection = "none",
  class = "display compact stripe",
  rownames = FALSE,
  options = list(
    dom = "frtip",
    pageLength = 25,
    lengthMenu = list(c(25, 50, 100, -1), c("25", "50", "100", "All")),
    scrollX = TRUE
  )
)
```

::: {.callout-edit}
Spot an inaccuracy or have a more recent guideline? Use the [contribute form](contribute.qmd) — submissions sync directly to the Google Sheet that this site reads from.
:::
