---
title: "*Plasmodium vivax* Policy Repository"
pagetitle: "P. vivax Policy Repository"
page-layout: full
toc: false
---

```{r}
#| label: load-data
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
})

source("R/data_helpers.R")

char <- load_characteristic_data()
case <- load_case_data()
stats <- headline_stats(char, case)
```

::: {.hero}
::: {.eyebrow}
A live policy repository
:::
# *Plasmodium vivax* Policy Repository

::: {.lede}
A continuously updated, country-level reference for *P. vivax* case management policies — treatment regimens, G6PD testing recommendations, implementation status, and the people behind the programmes. Built for national malaria control programmes, researchers, and partners working toward elimination.
:::

::: {.section-rule}
:::
:::

```{r}
#| output: asis

cat(sprintf(
  "
::: {.stat-row}
::: {.stat-card}
[%d]{.num}
[*P. vivax* endemic countries]{.label}
:::
::: {.stat-card}
[%d]{.num}
[Reporting *P. vivax* cases (last 5 yrs)]{.label}
:::
::: {.stat-card}
[%d]{.num}
[Have implemented G6PD testing]{.label}
:::
::: {.stat-card}
[%d]{.num}
[Have a documented policy date]{.label}
:::
:::
",
  stats$n_countries,
  stats$n_reporting,
  stats$n_implementing,
  stats$n_with_policy
))
```

## Why this exists

The tools available to test for and treat *P. vivax* malaria have changed rapidly in the last few years. Point-of-care G6PD tests, shorter and more effective radical cure regimens (single-dose tafenoquine, high-dose 7-day primaquine), and updated WHO global recommendations in 2023 and 2024 have triggered policy revisions across endemic countries.

National treatment guidelines are often published — but they are scattered, inconsistently versioned, and rarely available in a comparable form. A previous review was published in *Pathogens* in 2023. This repository is a living continuation of that effort, designed to stay current and to make every data point comparable across countries.

## What you can do here

```{=html}
<div class="country-grid">
  <a class="country-card" href="explore.html">
    <span class="name">Explore the data</span>
    <span class="meta">Filter by region, treatment, G6PD policy, programme phase, and more.</span>
  </a>
  <a class="country-card" href="countries.html">
    <span class="name">Country profiles</span>
    <span class="meta">One-page summaries of policy, implementation, and contacts.</span>
  </a>
  <a class="country-card" href="timeline.html">
    <span class="name">Policy timeline</span>
    <span class="meta">When countries last updated their guidelines and what they're considering next.</span>
  </a>
  <a class="country-card" href="contribute.html">
    <span class="name">Contribute an update</span>
    <span class="meta">National Malaria Control Programmes can submit changes via a short form.</span>
  </a>
</div>
```

::: {.callout-edit}
This repository is developed in consultation with National Malaria Control Programmes (NMCPs) and partners. NMCPs may flag specific data points as visible only to other programmes; the public site shows what countries have agreed to share publicly.
:::

## How it works

The data on this site is read at build time from a shared Google Sheet that is maintained by the repository team and updated by NMCP focal points. New submissions through the [contribute form](contribute.qmd) flow into the same sheet. When the site rebuilds (manually or on a schedule), the latest data is pulled and rendered. See [Methods](methods.qmd) for details, including how rows are reconciled and versioned.

## Cite this repository

If you use these data in a publication or report, please cite:

> *P. vivax* Policy Repository. Updated `r format(Sys.Date(), "%B %Y")`. Available at: https://example.org/vivax-policy-repository.

And the foundational review:

> Sadhewa A, Cassidy-Seyoum S, Acharya S, Devine A, Price RN, Mwaura M, et al. A Review of the Current Status of G6PD Deficiency Testing to Guide Radical Cure Treatment for Vivax Malaria. *Pathogens*. 2023;12(5):650.
