---
title: "Policy timeline"
subtitle: "When countries last updated their *P. vivax* case management guidelines, and what's coming next."
page-layout: full
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(jsonlite)
  library(stringr)
})

CASE_URL <- Sys.getenv("CASE_DATA_URL", "data/case_management.csv")
case <- read_csv(CASE_URL, show_col_types = FALSE, na = character()) |>
  mutate(across(everything(), as.character))
case[is.na(case)] <- ""
names(case) <- trimws(names(case))

case <- case |>
  filter(!str_detect(Country, " - "))

to_year <- function(v) {
  s <- str_trim(as.character(v))
  m <- str_extract(s, "(19|20)\\d{2}")
  out <- as.integer(m)
  out[!nzchar(s) | toupper(s) == "NA"] <- NA_integer_
  out
}

yr_df <- case |>
  mutate(
    year = to_year(`Last Policy Update`),
    Region = if_else(nzchar(Region), Region, "Unknown")
  ) |>
  filter(!is.na(year))

timeline_payload <- list(
  points = yr_df |>
    transmute(
      key = Country,
      country = Country,
      year = year,
      region = Region,
      drug = `Schizontocidal Drug`,
      `next` = `Next policy update (Y/N, Year)`
    ),
  meta = list(
    regionColors = list(
      Africa = "#c0521b",
      `Asia-Pacific` = "#0d4f4f",
      `Central and South America` = "#d9a36a",
      Unknown = "#8a9494"
    ),
    milestones = list(
      list(year = 2015, label = "WHO 2015 GL"),
      list(year = 2023, label = "WHO 2023 update"),
      list(year = 2024, label = "WHO 2024 update")
    )
  )
)
```

::: {.lede}
The last decade has seen rapid change in *P. vivax* policy. WHO global recommendations were updated in 2023 and 2024, triggering revisions in many countries. This page tracks when each country last formally revised its national treatment guidelines.
:::

## Last policy update by country

```{r}
#| label: timeline-data
#| echo: false
#| results: asis

cat(as.character(htmltools::tags$script(
  id = "timeline-data",
  type = "application/json",
  htmltools::HTML(jsonlite::toJSON(timeline_payload, auto_unbox = TRUE, dataframe = "rows", null = "null"))
)))
```

::: {.vivax-chart}
<div id="chart-timeline" class="vivax-chart__plot vivax-chart__plot--timeline"></div>
:::

## Countries with planned updates

```{r}
#| label: planned
#| output: asis

upcoming <- case |>
  filter(str_detect(`Next policy update (Y/N, Year)`, regex("^(yes|y,)", ignore_case = TRUE))) |>
  transmute(
    Country,
    `Planned update` = `Next policy update (Y/N, Year)`,
    `Treatments under consideration` = `Treatments Under Consideration`
  ) |>
  filter(nzchar(Country))

if (nrow(upcoming)) {
  knitr::kable(upcoming, format = "pipe")
} else {
  cat("_No countries have flagged a planned update at this time._")
}
```

## Key global moments

- **2015** — WHO recommends G6PD testing before primaquine in non-emergency settings.
- **2023** — WHO updates global malaria guidelines to include single-dose tafenoquine for *P. vivax* radical cure where G6PD status is known.
- **2024** — WHO further updates guidelines, expanding recommendations on shorter primaquine regimens.

Countries that updated their guidelines in 2023–2025 are the most likely to reflect the latest WHO recommendations. Those with guidelines older than 2018 may pre-date several major changes — see the [country profile](countries.qmd) for the specific version on file.
